---
title: "Sampling Means Drawn from the Exponential Distribution"
author: "Brian Finnegan"
date: "August 10, 2015"
output: 
  pdf_document: 
    fig_height: 3
    fig_width: 4
---
#Overview
The following reports on a simulation whereby repeated samples of means (n=40) drawn from the exponential distribtuion are shown to be approximately normal.

#Simulations
The probabilty density function of an exponential distribution is 
$$f(x, \lambda) = \lambda e^{-\lambda x}$$
for x ≥ 0, 0 for x < 0.

First, we draw 1,000 random observations from the exponential distribution with lambda = 0.2:
```{r}
set.seed(1)
lambda = .2
exp <- (rexp(1000, lambda))
```

While exp here is actually a sample from the exponential distribution, as opposed to the actual distribution itself, it is a very large sample, and visually we can see it approximates the distribution closely:

```{r echo=TRUE}
## Draw the histogram of the random sample from the exponential distribution
hist(exp, breaks=20, freq=F)
## Draw the actual distribution over the random sample
i = seq(0,40)
lines(i,dexp(i,lambda))
abline(v = mean(exp), col = "blue")

```

Next, we generate 1,000 instances of the mean of 40 random, iid exponentials for analysis of their means and variances relative of those of the underlying distribution (or more accurately the approximation thereof:
```{r }
set.seed(1)
mns = NULL
for(i in 1:1000) mns = c(mns, mean(rexp(40, lambda)))
```

#Sample Mean versus Theoretical Mean
The theoretical mean of an exponential distribution with lambda = .2 is 1/.2 or 5.  The mean of the approximation of the distribution, indicated by the blue line in the histogram above, is quite close to five.

Turning to the 1,000 samples of the mean of 40 exponentials, we see the smallest mean was 3.1, the largest 7.5 and the mean of the means, at 4.99 and represented in the histogram by the blue line, was very close to the theoretcial value of 5:
```{r echo=TRUE}

hist(mns, freq = f)
abline(v = mean(mns), col="blue")
```

This is because the sample mean is an unbiased estimator of the population mean.  While the mean of the samples' means is similar to that of the population, the distribution of those means is quite different from that of the underlying distribution.  

#Sample Variance versus Theoretical Variance

#Distribution
